---
title: "gapminder.Rmd"
author: "Wendy Erb"
date: "2/26/2021"
output: html_document
---

## Data
The Gapminder project was pioneered by Hans Rosling, who is famous for describing the prosperity of nations over time through famines, wars and other historic events with this beautiful data visualization in his 2006 TED Talk: The best stats you’ve ever seen.

Install libraries

```{r libraries, results = "hide"}
library(dslabs)
library(tidyverse)
```

View data

```{r setup}
view(gapminder)
```

```{r filter, echo=FALSE}
gap2011 <- gapminder %>% 
  as_tibble() %>% 
  filter(year == 2011)
```

## Life expectancy
Add some text and code chunks to plot patterns in the life_expectancy variable. 
Some ideas to explore:
What is the distribution of values (show e.g. in a histogram, a density plot, or boxplot (or all three!)) and does that vary based on continent?

```{r life.scatter, echo=FALSE}
ggplot(data = gap2011) +
  geom_point(mapping = aes(x = gdp, y = life_expectancy))
```


```{r life.hist, echo=FALSE}
ggplot(data = gap2011) +
  geom_histogram(mapping = aes(x = life_expectancy), binwidth=5)
```

```{r lifexcont.boxplot, echo=FALSE}
ggplot(data = gap2011) +
  geom_boxplot(mapping = aes(x = life_expectancy, y = continent))
```
```{r lifexcont.dens, echo=FALSE}
ggplot(data = gap2011) +
  geom_density(mapping = aes(x = life_expectancy)) +
  facet_wrap(~ continent)
```

How does life expectancy relate to infant mortality rates, fertility, population size, or GDP? How many variables can you display on the same plot and is the most effective or visually appealing way to display the patterns?

```{r lifexdemog.dens, echo=FALSE}
ggplot(gap2011, aes(x = life_expectancy, y = infant_mortality, color = fertility, size = population)) +
  geom_point()+
  geom_smooth(aes()) 
```


## Fertility
After you’ve done some exploration of life expectancy, move on to add some plots and text under your Fertility header.

```{r fertil.hist, echo=FALSE}
ggplot(data = gap2011) +
  geom_histogram(mapping = aes(x = fertility), binwidth=0.5)
```
```{r fertxcont.dens, echo=FALSE}
ggplot(data = gap2011) +
  geom_density(mapping = aes(x = fertility)) +
  facet_wrap(~ continent)
```

## Infant mortality

```{r life.plot, echo=FALSE}
ggplot(data = gap2011) +
  geom_point(mapping = aes(x = gdp, y = life_expectancy))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
